---
title: "RepResearch Project 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# libraries
library(ggplot2)
library(dplyr)


# 1. Import data
library(readr)
activity <- read_csv("C:/Users/ANAJM/OneDrive - CIMMYT/Desktop/Reproducible research Assignment 2/activity.csv")
View(activity)

# 2. Process/transform the data (if necessary) into a format suitable for your analysis
```{r echo=TRUE}
A <- activity
A$date <- as.Date(A$date, format = "%Y-%m-%d")
```
# Questions and Solutions
### What is mean total number of steps taken per day?
**1. Total number of steps taken per day**
```{r echo=TRUE}
stepsperday <- aggregate(steps ~ date, data = A, FUN = "sum", na.rm = TRUE)
head(stepsperday)
```
**2. Histogram of total number of steps taken per day**
```{r}
hist(stepsperday$steps, xlab = "Number of steps per day", main ="Histogram of total steps taken per day", col = "grey")
```

**3. Mean and Median of the total number of steps taken per day**
```{r}
mean <- format(mean(stepsperday$steps), 2)
mean
median <- format(median(stepsperday$steps), 2)
median
```
* Mean: **`r mean`**
* Median: **`r median`**

### Average daily activity pattern

**1. Time series plot of the 5-minute interval (x-axis) and average number of steps taken (y-axis)**
```{r}
int <- aggregate(steps ~ interval, data=A, FUN = "mean", na.rm=TRUE)
with(int, plot(interval, steps, type = "l", xlab = "Interval", ylab = "Number of steps", main = "Average steps per interval", col="red"))
```

**2. 5-minute interval on average containing maximum number of steps**
```{r}
indexmax <- which(int$steps==max(int$steps))
maxsteps <- int$steps[indexmax]
maxinterval <- int$interval[indexmax]
```
* Highest average number of steps per day: **`r maxsteps`**
* Interval with highest average number of steps: **`r maxinterval`**

### Impute Missing values 
**1. Calculate and report the total number of missing values in the dataset**
```{r}
totna <- sum(is.na(A$steps))
totna
```
* Total number of rows with NA's: **`r totna`**

**2. Filling in all of the missing values in the dataset.**
**3. Create a new dataset with the missing data filled in.**
```{r}
Aimputed <- na.mean(A)
head(Aimputed)
```

**4. Histogram of total number of steps taken each day. Calculation of  mean and median total number of steps taken per day. 
```{r}
stepsnew <- aggregate(steps ~ date, data = Aimputed, FUN = "sum", na.rm = TRUE)
hist(stepsnew$steps, xlab = "Number of steps per day", main ="Histogram of total steps taken per day (NA's replaced)", col = "blue")
meannew <- format(mean(stepsnew$steps))
meannew
mediannew <- format(median(stepsnew$steps), 2)
mediannew
```
Diff between the original mean/median and the new mean/median (NA's replaced with the mean)

* Original Mean: **`r mean`**
* New Mean (NA's replaced): **`r meannew`**

* Original Median: **`r median`**
* New Median (NA's replaced): **`r mediannew`**

### Differences in activity patterns between weekdays and weekends?
**1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" **
```{r}
Aimputed$daytype <- ifelse(weekdays(Aimputed$date)== "Sunday" | weekdays(Aimputed$date)== "Saturday", "weekend", "weekdays")
head(Aimputed)
```

**2. Panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken (y-axis).**
```{r}
library(lattice)
stepsimputed <- aggregate(steps ~ interval + daytype, data=Aimputed, FUN = "mean", na.rm=TRUE)
xyplot(stepsimputed$steps ~ stepsimputed$interval | daytype, data = stepsimputed, type="l", ylab = "Number of steps", main="Activity pattern per weekday/weekend", xlab = "Interval", layout=c(1,2))
```